---
title: "Women of 2020"
author: "Mike Wehinger"
date: "08/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
#library(scales)
library(kableExtra)

#women <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-12-08/women.csv')

women <- readr::read_csv('women.csv')



```

## A Sunburst Chart

The BBC has revealed its list of 100 inspiring and influential women from around the world for 2020. For each woman, we know the category assigned by the BBC and their role/career. 

A sunburst chart will break-down the list of 100 by category and then by role

```{r explore, message=FALSE}

head(women)

kable(table(women$category), caption = "Cateogries")

kable(table(women[,c("role", "category")]), caption = "Cateogries and Roles")

```

## Cleaning Roles

Some roles can be cleaned, for example, the following roles can be classified as **Scientist**

* Scientist	
* Scientist and inventor	
* Scientist/astronaut

```{r roles, echo=TRUE}

women$role <- str_to_lower(women$role, locale = "en")

women <- women %>% mutate(cleanRole = case_when(
  str_detect(role, "activist") ~ "Activist",
  str_detect(role, "scientist") ~ "Scientist",
  str_detect(role, "writer") ~ "Writer",
  str_detect(role, "blogger") ~ "Writer",
  str_detect(role, "campaigner") ~ "Campaigner",
  str_detect(role, "protest leader") ~ "Campaigner",
  str_detect(role, "actor") ~ "Actress",
  str_detect(role, "actress") ~ "Actress",
  str_detect(role, "health") ~ "Health Expert",
  str_detect(role, "pathologist") ~ "Health Expert",
  str_detect(role, "nutritionist") ~ "Health Expert",
  str_detect(role, "nurse") ~ "Health Expert",
  str_detect(role, "doctor") ~ "Health Expert",
  str_detect(role, "minister") ~ "Politician",
  str_detect(role, "mayor") ~ "Politician",
  str_detect(role, "parliamentary") ~ "Politician",
  str_detect(role, "senator") ~ "Politician",
  str_detect(role, "deputy") ~ "Politician",
  str_detect(role, "ambassador") ~ "Politician",
  str_detect(role, "artist") ~ "Artist",
  str_detect(role, "musician") ~ "Artist",
  str_detect(role, "poet") ~ "Artist",
  str_detect(role, "rapper") ~ "Artist",
  str_detect(role, "singer") ~ "Artist",
  str_detect(role, "model") ~ "Artist",
  TRUE ~ role))

women$cleanRole <- str_to_title(women$cleanRole, locale = "en")

#kable(table(women$role))

#kable(table(women$cleanRole))

kable(table(women[,c("cleanRole", "category")]), caption = "Cateogries and Roles")


```


## Creating the Chart

The chart will be made with plotly because plotly has a handy sunburst trace. The trick is to get the data into the correct format.

```{r tryPlotly, echo=TRUE, message=FALSE}
library(plotly)

# Prepare the data for the sunburst chart. 

# 1. group and count by category and role and rename the cols to identify the components needed for the chart
plotlyData <- women %>% group_by(category, cleanRole) %>% summarize(count = n()) %>% arrange(desc(count)) %>% rename(parents = category, labels = cleanRole)

# 2. Add just the category counts
plotlyData <- plotlyData %>% bind_rows(women %>% group_by(category) %>% summarize(count = n()) %>% arrange(count) %>% rename(labels = category) %>% mutate(parents = ""))

# 3. Initiate an ID field and populate it it based on the parent and labels values 
plotlyData$ids <- ""

plotlyData[plotlyData$parents!="",]$ids <- paste(plotlyData[plotlyData$parents!="",]$parents, plotlyData[plotlyData$parents!="",]$labels, sep="-")

plotlyData[plotlyData$parents=="",]$ids <- plotlyData[plotlyData$parents=="",]$labels

plotlyChart <- plot_ly(plotlyData, ids = ~ids, labels = ~labels, parents = ~parents, values = ~count, type = 'sunburst', branchvalues="total", maxdepth=2,insidetextorientation='radial')

plotlyChart

```


```{r chart, echo=TRUE, eval=FALSE}
foundation <- women %>% summarize(count=n())

sunburst_0 = ggplot(foundation) + 
  geom_bar(data=foundation, aes(x=1, y=count), fill='darkgrey', stat='identity') +
  geom_text(aes(x=1, y=count, label="100 Influential Women in 2020"), color='white')

sunburst_0 + coord_polar("y")

first <- women %>% group_by(category) %>% summarize(count = n()) %>% arrange(desc(count))

sunburst_1 <- sunburst_0 + 
  geom_bar(data=first, aes(x=2, y=count, fill=count),
 color='white', position='stack', stat='identity', size=0.6) + 
 geom_text(data=first, aes(label=paste(category, count), x=2, y=count), position='stack')

sunburst_1

second <- women %>% group_by(category, cleanRole) %>% summarize(count = n()) %>% arrange(desc(count))

sunburst_2 <- sunburst_1 + 
  geom_bar(data=second, aes(x=3, y=count, fill=count),
 color='white', position='stack', stat='identity', size=0.6) + 
 geom_text(data=second, aes(label=paste(cleanRole, count), x=3, y=count), position='stack')

sunburst_2


#sunburst_1 + coord_polar("y")
```

