---
title: "IKEA Analysis"
author: "Mike Wehinger"
date: "03/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotly)

ikea <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-11-03/ikea.csv')
```

Some data cleaning from the initial exploratory analysis. 

```{r cleaning, include=TRUE, warning=FALSE, message=FALSE}
# Clean up the Designer column
## Remove 'IKEA of Sweden/' and '/IKEA of Sweden'
ikea$designer <- ikea$designer %>% str_remove("/IKEA of Sweden")
ikea$designer <- ikea$designer %>% str_remove("IKEA of Sweden/")

## Remove designers that have a character length over 39 (they are not names)
ikea <- ikea %>% mutate(len = nchar(designer)) %>% filter(len < 40)

## Remove rows that have an NA in either width, depth, or height...
ikea <- ikea %>% drop_na(c("width", "height", "depth"))

## Get the top x designers by product count
top <- 10
topDesigners <- ikea %>% group_by(designer) %>% summarize(n=n()) %>% arrange(desc(n)) %>% slice_head(n=top)

## Add top 20 designers to the main table. 
ikea <- ikea %>% left_join(topDesigners, by="designer")# %>% filter(n>0)

```

Using the product dimensions (width, height, depth), perform a k-means cluster analysis to see what are the common grouped sizes of products. Trying group sizes from three to six and visualizaing with a 3d scatterplot

```{r clustering, include=TRUE, warning=FALSE, message=FALSE}
## Find clusters in product dimensions using k-means...
ikea <- ikea %>% mutate(widthS = scale(width), heightS = scale(height), depthS=scale(depth))

ikea$cluster3 <- kmeans(ikea[,c("widthS", "heightS", "depthS")],3)$cluster
ikea$cluster4 <- kmeans(ikea[,c("widthS", "heightS", "depthS")],4)$cluster
ikea$cluster5 <- kmeans(ikea[,c("widthS", "heightS", "depthS")],5)$cluster
ikea$cluster6 <- kmeans(ikea[,c("widthS", "heightS", "depthS")],6)$cluster

ikea %>% plot_ly(x=~width, y=~depth, z=~height, color=as.factor(ikea$cluster4), marker=list(symbol = "square"))

table(ikea[,c("category", "cluster4")])

```

Calculate product size by determining the how many cubic meters the piece is (simply width * height * depth). Compare this to price. For example, it could be that the larger the furniture, the more material and handling costs, hence the higher the price. 

Also, add average size line for each cluster

```{r sizePlots, warning=FALSE, message=FALSE}
ikea <- ikea %>% mutate(size = (width/100)*(height/100)*(depth/100))

ClusterAvg <- ikea %>% group_by(cluster4) %>% summarize(a=mean(size))

ikea %>% plot_ly(x = ~size, y = ~price, color = as.factor(ikea$cluster4),
                 marker = list(symbol="square", size = 7,
                               line = list(color = 'rgba(152, 0, 0, .95)', width = 1))) %>% 
    layout(shapes=list(list(type='line', x0= ClusterAvg[1,]$a, x1= ClusterAvg[1,]$a, y0=min(ikea$price), y1=max(ikea$price), line=list(dash='dot', width=1)), list(type='line', x0= ClusterAvg[2,]$a, x1= ClusterAvg[2,]$a, y0=min(ikea$price), y1=max(ikea$price), line=list(dash='dot', width=1)), list(type='line', x0= ClusterAvg[3,]$a, x1= ClusterAvg[3,]$a, y0=min(ikea$price), y1=max(ikea$price), line=list(dash='dot', width=1)), list(type='line', x0= ClusterAvg[4,]$a, x1= ClusterAvg[4,]$a, y0=min(ikea$price), y1=max(ikea$price), line=list(dash='dot', width=1))), legend = list(x=.75, y=1,bgcolor = "#E2E2E2", title=list(text='Size Category')),title = "IKEA Furniture Size vs Price", yaxis = list(zeroline = TRUE, title="Price (SAR)"),xaxis = list(zeroline = TRUE, title="Size (Cubic Meters)"))


```

The average size lines make the plot too busy... How about looking at product category by size and price, 

```{r CatPlots, warning=FALSE, message=FALSE}
# Make plot,

catPlot2 <- ggplot(ikea,aes(x=price, y=category, colour=size)) +
    geom_jitter(aes(text=paste("Name: ", name)), width=0.25, alpha=0.5, ) +
    labs(title = "Product Category by price",
         x = "Price",
         y = "Category")

    ggplotly(catPlot2)
```

The scatter plot is better, just needs to be cleaned up a bit... 

```{r ScatPlots, warning=FALSE, message=FALSE}
ikea <- ikea %>% mutate(clusterName = 
                  case_when(
                    cluster4 == 1 ~ "Tall (e.g. wardrobes/Bookcase)", 
                    cluster4 == 2 ~ "Wide (e.g. sofas/beds)", 
                    cluster4 == 3 ~ "Deep (e.g. chairs/tables)", 
                    cluster4 == 4 ~ "Smaller (e.g. TVs)", TRUE ~ "" ))

pal <- c("#779761",  "#A85105", "#055CA8", "#F5D817")    

sizePlot2 <- ikea %>% plot_ly(x = ~size, y = ~price, color = as.factor(ikea$clusterName), colors=pal,
                 text = paste("Item: ", ikea$name, "<br>Category :", ikea$category, "<br>Price:", ikea$price, "<br>Width: ", ikea$width,"<br>Height: ", ikea$height, "<br>Depth: ", ikea$depth, "<br>Shape:", ikea$clusterName),
                 marker = list(symbol="square", size = 7, line = list(color = pal, width = .5))) %>% 
    layout(
        legend = list(bgcolor = "#FFFFFF", title=list(text='Size Category')),
        title = "IKEA Furniture Size vs Price", 
        yaxis = list(zeroline = TRUE, title="Price (SAR)"),
        xaxis = list(showgrid = FALSE, zeroline = TRUE, title="Size (Cubic Meters)"))

sizePlot2

            
```

