---
title: "IKEA Random Forest Tuning"
author: "Mike Wehinger"
date: "13/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, cache = TRUE)
library(tidyverse)
library(tidymodels)
theme_set(theme_minimal())

ikea <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-11-03/ikea.csv')

```
Walking through the [Julia Silge video](https://juliasilge.com/blog/ikea-prices/) about tuning random forests using the Tidy Tuesday IKEA dataset.  

## Exploratory Analysis

Visualizing the data with price as the Y and width, height, and depth as Xs. Take logs of price due to distribution. 

```{r}
ikea %>% select(X1, price, depth:width) %>% 
  pivot_longer(depth:width, names_to = "dim") %>% 
  ggplot(aes(value, price, color=dim)) + 
  geom_point(alpha=.4, show.legend = FALSE) +
  scale_y_log10() + 
  facet_wrap(~dim, scales = "free_x") +
  labs(x=NULL)
  
```


```{r}

ikea_df <- ikea %>% select(price, name, category, depth, height, width) %>%
  mutate(price = log10(price)) %>% 
  mutate_if(is.character, factor)

```

## Build Model

Use tidymodels to split data into training/testing based on a price-stratified random sample  

Then use bootstraps to create 25 subsets on the training data

```{r}
set.seed(123)
ikea_split <- initial_split(ikea_df, strata = price)
ikea_train <- training(ikea_split)
ikea_test <- testing(ikea_split)

set.seed(234)
ikea_folds <- bootstraps(ikea_train, strata = price)
ikea_folds

```

The usemodels package is a cool set of model creating functions. It will generate the code needed to stand up common models.  

In this case, you can set up a ranger random forest model using *use_ranger()* function - see the output...

```{r}
library(usemodels)

use_ranger(price ~ ., data= ikea_train)


```

Using the above function, we cut and paste the code and then modify to suit your purposes. 

```{r}

library(textrecipes)

ranger_recipe <- 
  recipe(formula = price ~ ., data = ikea_train) %>% 
  step_other(name, category, threshold = .01) %>%
  step_clean_levels(name, category) %>%
  step_knnimpute(depth, height, width)

ranger_spec <- 
  rand_forest(mtry = tune(), min_n = tune(), trees = 1000) %>% 
  set_mode("regression") %>% 
  set_engine("ranger") 

ranger_workflow <- 
  workflow() %>% 
  add_recipe(ranger_recipe) %>% 
  add_model(ranger_spec) 

set.seed(91010)

ranger_tune <-
  tune_grid(ranger_workflow, 
            resamples = ikea_folds, 
            grid = 11)


```


## Explore Results

Look at metrics from the possible models...

```{r}
show_best(ranger_tune, metric="rmse")
show_best(ranger_tune, metric="rsq")

autoplot(ranger_tune)


```

Create finalized workflow using the best parameters 

```{r}

final_rf <- ranger_workflow %>%
  finalize_workflow(select_best(ranger_tune))

final_rf

```

Create the final fit with the training and testing data... 

```{r}

ikea_fit <- last_fit(final_rf, ikea_split)
ikea_fit


```

Look at the metrics computed on the text set to compare against the re-sample results (above) 

```{r}
collect_metrics(ikea_fit)
```

Collect the predictions on the test set and plot the true vs predicted price from the test set. 


```{r}
collect_predictions(ikea_fit) %>%
  ggplot(aes(price, .pred)) +
  geom_abline(lty = 2, color="gray50") +
  geom_point(alpha = .5, color = "midnightblue") + 
  coord_fixed()

```
The trained workflow is in the ikea_fits object *ikea_fit$.workflow[[1]]* and if you have data, for example, row 15 of the test set, use it to predict a price

```{r}
predict(ikea_fit$.workflow[[1]], ikea_test[15,])
ikea_test[15,]

predict(ikea_fit$.workflow[[1]], ikea_test[150,])
ikea_test[150,]

predict(ikea_fit$.workflow[[1]], ikea_test[100,])
ikea_test[100,]
```
## Exploration of Feature Importance

Width is the most important thing and name is the least important. 

```{r}
library(vip)

imp_spec <- ranger_spec %>% 
  finalize_model(select_best(ranger_tune)) %>%
  set_engine("ranger", importance = "permutation")

workflow() %>%
  add_recipe(ranger_recipe) %>%
  add_model(imp_spec) %>%
  fit(ikea_train) %>%
  pull_workflow_fit() %>%
  vip(aesthetics = list(alpha=.8, fill = "midnightblue"))


```

