---
title: "SuperBowlAds"
author: "Mike Wehinger"
date: "03/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

#youtube <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-03-02/youtube.csv')

youtube <- readr::read_csv('youtube.csv')

```

## Super Bowl Ad Analysis

This data comes from [FiveThirtyEight](https://github.com/fivethirtyeight/superbowl-ads) where they classified 233 ads from 10 brands that aired the most spots in all 21 Super Bowls this century using seven specific criteria. 

This analysis will use that data to see what strategies were used by year and/or brand considering youtube views

### Prepare the Data

The following code transforms the data into a tidy list of strategies...

```{r DataPrep, echo=FALSE}

# In the view_count data replace NAs with a zero
youtube$view_count <- youtube$view_count %>% replace_na(0)

# For each ad, convert the strategy TRUE/FALSE indicator with a 1 or 0...

youtube <- youtube %>% mutate(funny = if_else(funny==TRUE,1,0)) %>%
  mutate(show_product_quickly = if_else(show_product_quickly==TRUE,1,0)) %>%
  mutate(patriotic = if_else(patriotic==TRUE,1,0)) %>%
  mutate(celebrity = if_else(celebrity==TRUE,1,0)) %>%
  mutate(danger = if_else(danger==TRUE,1,0)) %>%
  mutate(animals = if_else(animals==TRUE,1,0)) %>%
  mutate(use_sex = if_else(use_sex==TRUE,1,0))

# Count the strategies used in each add...
youtube$Strategy_Count <- youtube$funny + youtube$patriotic + youtube$celebrity + youtube$danger + youtube$animals + youtube$use_sex

# Add a strategy category of none and select only the cols of interest and ignore ads with no strategy
youtube <- youtube %>% mutate(No_Strategy = ifelse(Strategy_Count==0,1,0)) %>% 
  filter(Strategy_Count>0) %>%
  select(year,brand,funny,patriotic,celebrity,danger,animals, use_sex, Strategy_Count, view_count)

# Clean Up column names
names(youtube) <- gsub("_", " ", names(youtube)) %>% str_to_title()

# Pivot the table long such that each strategy for an ad comprises a row. For example, an ad that used three
# strategies will have three rows.
youtube <- youtube %>% pivot_longer(Funny:'Use Sex', names_to="Strategy", values_to = "Count")

# Then allocate the strategy count evenly across the strategies. For example, an ad using three strategies = each strategy will have been used .333 times in that ad.
youtube$`Strategy Count` <- youtube$Count / youtube$`Strategy Count`

# Allocate the view account evenly across the strategies too, 
youtube$`View Count` <- youtube$`View Count` * youtube$`Strategy Count`

# Filter out rows of unused strategies..
youtube <- youtube %>% filter(Count > 0) %>% select(-Count)

# Now the data is prepared and can be grouped according to vizual requirements 
```



### Visualizations

Lets try a Mosaic plot where the bar width is determined by youtube Views and the height is a 100% stacked bar showing proportion of strategies used...

```{r MosaicChart, echo=FALSE}
#sub2 <- youtube %>% drop_na %>% group_by(year, use_sex) %>% summarize(count = n(), Views = sum(view_count)) %>% mutate(year.count=sum(count), prop = count/sum(count)) %>% ungroup()

# A Mosaic plot looking by year...
print("Mosaic By Year")
youtube %>% group_by(Year, Strategy) %>% summarize(`Strategy Count` =sum(`Strategy Count`), `View Count` = sum(`View Count`)) %>% mutate(view_count = sum(`View Count`), prop = `View Count` / sum(`View Count`), 'Ad Count'= sum(`Strategy Count`)) %>% ungroup() %>% ggplot(aes(x = `Ad Count`, y = prop, width = `Ad Count`, fill = Strategy)) +
    geom_bar(stat = "identity", position = "fill", colour = "black") +
    # geom_text(aes(label = scales::percent(prop)), position = position_stack(vjust = 0.5)) + # if labels are desired
    facet_grid(~Year, scales = "free_x", space = "free_x") +
    scale_fill_brewer(palette = "RdYlGn")


print("Mosaic for 2019 & 2020")
youtube %>% filter(Year > 2018) %>% group_by(Year, Strategy) %>% summarize(`Strategy Count` =sum(`Strategy Count`), `View Count` = sum(`View Count`)) %>% mutate(strat_count = sum(`Strategy Count`), prop = `Strategy Count` / sum(`Strategy Count`), Views = sum(`View Count`)) %>% ungroup() %>% ggplot(aes(x = strat_count, y = prop, width = Views, fill = Strategy)) +
    geom_bar(stat = "identity", position = "fill", colour = "black") +
    # geom_text(aes(label = scales::percent(prop)), position = position_stack(vjust = 0.5)) + # if labels are desired
    facet_grid(~Year, scales = "free_x", space = "free_x") +
    scale_fill_brewer(palette = "RdYlGn")




# This time by brand...
print("Mosaic By Brand")
youtube %>% group_by(Brand, Strategy) %>% summarize(`Strategy Count` =sum(`Strategy Count`), `View Count` = sum(`View Count`)) %>% mutate(strat_count = sum(`Strategy Count`), prop = `Strategy Count` / sum(`Strategy Count`), Views = sum(`View Count`)) %>% ungroup() %>% ggplot(aes(x = strat_count, y = prop, width = Views, fill = Strategy)) +
    geom_bar(stat = "identity", position = "fill", colour = "black") +
    # geom_text(aes(label = scales::percent(prop)), position = position_stack(vjust = 0.5)) + # if labels are desired
    facet_grid(~Brand, scales = "free_x", space = "free_x") +
    scale_fill_brewer(palette = "RdYlGn")


print("Mosaic By Year for Budweiser")
youtube %>% filter(Brand == "Budweiser") %>% group_by(Year, Strategy) %>% summarize(`Strategy Count` =sum(`Strategy Count`), `View Count` = sum(`View Count`)) %>% mutate(strat_count = sum(`Strategy Count`), prop = `Strategy Count` / sum(`Strategy Count`), Views = sum(`View Count`)) %>% ungroup() %>% ggplot(aes(x = strat_count, y = prop, width = Views, fill = Strategy)) +
    geom_bar(stat = "identity", position = "fill", colour = "black") +
    # geom_text(aes(label = scales::percent(prop)), position = position_stack(vjust = 0.5)) + # if labels are desired
    facet_grid(~Year, scales = "free_x", space = "free_x") +
    scale_fill_brewer(palette = "RdYlGn")


print("Mosaic By Brand for 2020")
youtube %>% filter(Year==2020) %>% group_by(Brand, Strategy) %>% summarize(`Strategy Count` =sum(`Strategy Count`), `View Count` = sum(`View Count`)) %>% mutate(strat_count = sum(`Strategy Count`), prop = `Strategy Count` / sum(`Strategy Count`), Views = sum(`View Count`)) %>% ungroup() %>% ggplot(aes(x = strat_count, y = prop, width = Views, fill = Strategy)) +
    geom_bar(stat = "identity", position = "fill", colour = "black") +
    # geom_text(aes(label = scales::percent(prop)), position = position_stack(vjust = 0.5)) + # if labels are desired
    facet_grid(~Brand, scales = "free_x", space = "free_x") +
    scale_fill_brewer(palette = "RdYlGn")


# Both (excluding brand)
print("Mosaic By Year and Brand (excluding Doritos)")
youtube %>% filter(Brand != "Doritos") %>% group_by(Year, Brand, Strategy) %>% summarize(`Strategy Count` =sum(`Strategy Count`), `View Count` = sum(`View Count`)) %>% mutate(strat_count = sum(`Strategy Count`), prop = `Strategy Count` / sum(`Strategy Count`), Views = sum(`View Count`)) %>% ungroup() %>% ggplot(aes(x = strat_count, y = prop, width = Views, fill = Strategy)) +
    geom_bar(stat = "identity", position = "fill", colour = "black") +
    # geom_text(aes(label = scales::percent(prop)), position = position_stack(vjust = 0.5)) + # if labels are desired
    facet_grid(Brand~Year, scales = "free_x", space = "free_x") +
    scale_fill_brewer(palette = "RdYlGn") + theme_void() 

    # theme(panel.spacing.x = unit(0, "npc")) + # if no spacing preferred between bars



```

Having the youtube view count on the x-axis is distorting the viz too much. Try moving it to the Y axis, which suggests each strategy is responsible for a certain proportion of views. The x-axis can be add count, which is usually 8 to 12

```{r MosaicChartv2, echo=FALSE}

print("Mosaic By Year")
youtube %>% group_by(Year, Strategy) %>% summarize(`Strategy Count` =sum(`Strategy Count`), `View Count` = sum(`View Count`)) %>% mutate(strat_count = sum(`Strategy Count`), prop = `Strategy Count` / sum(`Strategy Count`), Views = sum(`View Count`)) %>% ungroup() %>% ggplot(aes(x = strat_count, y = prop, width = Views, fill = Strategy)) +
    geom_bar(stat = "identity", position = "fill", colour = "black") +
    # geom_text(aes(label = scales::percent(prop)), position = position_stack(vjust = 0.5)) + # if labels are desired
    facet_grid(~Year, scales = "free_x", space = "free_x") +
    scale_fill_brewer(palette = "RdYlGn")

print("Mosaic By Brand")
youtube %>% group_by(Brand, Strategy) %>% summarize(`Strategy Count` =sum(`Strategy Count`), `View Count` = sum(`View Count`)) %>% mutate(view_count = sum(`View Count`), prop = `View Count` / sum(`View Count`), 'Ad Count'= sum(`Strategy Count`)) %>% ungroup() %>% ggplot(aes(x = `Ad Count`, y = prop, width = `Ad Count`, fill = Strategy)) +
    geom_bar(stat = "identity", position = "fill", colour = "black") +
    # geom_text(aes(label = scales::percent(prop)), position = position_stack(vjust = 0.5)) + # if labels are desired
    facet_grid(~Brand, scales = "free_x", space = "free_x") +
    scale_fill_brewer(palette = "RdYlGn")
```


This might as well be a stacked bar chart - taking away the info provided by the width. Maybe try a heatmap.

```{r Heat, echo=FALSE}
print("Heatmap By Year")
youtube %>% group_by(Year, Strategy) %>% summarize(`Strategy Count` =sum(`Strategy Count`), `View Count` = sum(`View Count`)) %>% mutate(view_count = sum(`View Count`), prop = `View Count` / sum(`View Count`), 'Ad Count'= sum(`Strategy Count`)) %>% ungroup()%>% ggplot(aes(Year, Strategy, fill=prop)) + 
    geom_tile(color="white") + scale_x_discrete()+ coord_equal() + theme(axis.ticks.y = element_blank(),panel.background = element_blank(),legend.position = c(0.87, 1.21),legend.background = element_blank(),legend.direction = "horizontal")+scale_fill_continuous(high = "#132B43", low = "#56B1F7")

print("Heatmap By Brand")
youtube %>% group_by(Brand, Strategy) %>% summarize(`Strategy Count` =sum(`Strategy Count`), `View Count` = sum(`View Count`)) %>% mutate(view_count = sum(`View Count`), prop = `View Count` / sum(`View Count`), 'Ad Count'= sum(`Strategy Count`)) %>% ungroup()%>% ggplot(aes(Brand, Strategy, fill=prop)) + 
    geom_tile(color="white") + scale_x_discrete()+ coord_equal() + theme(axis.ticks.y = element_blank(),panel.background = element_blank(),legend.position = c(0.87, 1.21),legend.background = element_blank(),legend.direction = "horizontal")+scale_fill_continuous(high = "#132B43", low = "#56B1F7")

```

Maybe that is not too interesting though, try looking out youtube view counts relating to combinations of strategies...

```{r multiStrat, eval=TRUE}

youtube <- readr::read_csv('youtube.csv')

youtube$view_count <- youtube$view_count %>% replace_na(0)

youtube$strategy <- paste(
  if_else(youtube$funny==TRUE, "Funny", ""),
  if_else(youtube$patriotic==TRUE, "Patriotic", ""),
  if_else(youtube$celebrity==TRUE, "Celebrity", ""),
  if_else(youtube$danger==TRUE, "Danger", ""),
  if_else(youtube$animals==TRUE, "Animals", ""),
  if_else(youtube$use_sex==TRUE, "Sex", ""), sep=" ")

youtube$strategy <- youtube$strategy %>% str_squish()

youtube$strategy <- if_else(youtube$strategy=="", "None", youtube$strategy)

youtube <- youtube %>% select(year,brand,strategy, view_count) %>% arrange(desc(view_count))

youtubeHeat <- youtube %>% group_by(year, strategy) %>% summarize(Views = sum(view_count)) %>% mutate(view_count = sum(Views), percentViews = Views /sum(Views))

youtubeSort <- youtube %>% group_by(strategy) %>% summarize(Views = sum(view_count)) %>% mutate(view_count = sum(Views), percentViews = Views /sum(Views))

youtubeSort$strategy <- factor(youtubeSort$strategy, levels = youtubeSort$strategy[order(youtubeSort$Views)])

youtubeSort %>% ggplot(aes(strategy, Views)) +                                    # Decreasingly ordered barchart
    geom_bar(stat = "identity") + coord_flip()

youtubeSort %>% filter(Views > 750000) %>% ggplot(aes(strategy, Views)) +                                    # Decreasingly ordered barchart
    geom_bar(stat = "identity") + coord_flip()

```

How about bringing in the likes and dislikes in addition to the total views...

```{r Engagement, eval=TRUE}

youtube <- readr::read_csv('youtube.csv')

youtube$view_count <- youtube$view_count %>% replace_na(0)

youtube$strategy <- paste(
  if_else(youtube$funny==TRUE, "Funny", ""),
  if_else(youtube$patriotic==TRUE, "Patriotic", ""),
  if_else(youtube$celebrity==TRUE, "Celebrity", ""),
  if_else(youtube$danger==TRUE, "Danger", ""),
  if_else(youtube$animals==TRUE, "Animals", ""),
  if_else(youtube$use_sex==TRUE, "Sex", ""), sep=" ")

youtube$strategy <- youtube$strategy %>% str_squish()

youtube$strategy <- if_else(youtube$strategy=="", "None", youtube$strategy)

youtube <- youtube <- youtube %>% select(year,brand,strategy, view_count, like_count, dislike_count,comment_count) %>% arrange(desc(view_count))

youtube$dislike_count <- youtube$dislike_count * -1 

tmp <- youtube %>% group_by(strategy) %>% summarize(Views = sum(view_count), Likes = sum(like_count), Dislike = sum(dislike_count), Comment = sum(comment_count)) %>% pivot_longer(Views:Comment, names_to = "stat", values_to = "Count")

tmp$Count <- tmp$Count %>% replace_na(0)

#tmp$strategy <- factor(tmp$strategy, levels = tmp$strategy[order(tmp$Views)])

tmp$strategy <- factor(tmp$strategy, levels = tmp[tmp$stat=="Views",]$strategy[order(tmp[tmp$stat=="Views",]$Count)])

tmp %>% ggplot(aes(x=strategy, y=Count, fill=stat)) + 
  geom_bar(data = subset(tmp, stat=="Likes"), stat="identity") + 
  geom_bar(data = subset(tmp, stat=="Dislike"), stat="identity") + 
  coord_flip()

tmp %>% ggplot(aes(x=strategy, y=Count, fill=stat)) + 
    geom_bar(data = subset(tmp, stat=="Likes"), stat="identity") + 
    geom_bar(data = subset(tmp, stat=="Dislike"), stat="identity") + geom_point(data = subset(tmp, stat=="Comment")) +
    coord_flip()

```


```{r oldApproach, eval=FALSE}

youtube <- readr::read_csv('youtube.csv')

youtube$view_count <- youtube$view_count %>% replace_na(0)

youtube$strategy <- paste(
  if_else(youtube$funny==TRUE, "Funny", ""),
  if_else(youtube$patriotic==TRUE, "Patriotic", ""),
  if_else(youtube$celebrity==TRUE, "Celebrity", ""),
  if_else(youtube$danger==TRUE, "Danger", ""),
  if_else(youtube$animals==TRUE, "Animals", ""),
  if_else(youtube$use_sex==TRUE, "Sex", ""), sep=" ")

youtube$strategy <- youtube$strategy %>% str_squish()

youtube$strategy <- if_else(youtube$strategy=="", "None", youtube$strategy)


sub <- youtube %>% drop_na() %>% group_by(year, brand, strategy) %>% summarize(Views = sum(view_count))



sub_long <- youtube %>% mutate(funny = if_else(funny==TRUE,1,0)) %>% mutate(show_product_quickly = if_else(show_product_quickly==TRUE,1,0))%>% mutate(patriotic = if_else(patriotic==TRUE,1,0))%>% mutate(celebrity = if_else(celebrity==TRUE,1,0))%>% mutate(danger = if_else(danger==TRUE,1,0))%>% mutate(animals = if_else(animals==TRUE,1,0))%>% mutate(use_sex = if_else(use_sex==TRUE,1,0)) %>% select(year,brand, funny, show_product_quickly, patriotic, celebrity, danger, animals, use_sex)

sub_long %>% group_by(year, Strategy) %>% summarize(Count=sum(Count))

sub_long2 <- sub_long %>% group_by(year, Strategy) %>% summarize(Count=sum(Count)) %>% mutate(year.count = sum(Count), prop = Count/sum(Count))

ggplot(sub_long2,
       aes(x = year, y = prop, width = year.count, fill = Strategy)) +
    geom_bar(stat = "identity", position = "fill", colour = "black") +
    # geom_text(aes(label = scales::percent(prop)), position = position_stack(vjust = 0.5)) + # if labels are desired
    facet_grid(~year, scales = "free_x", space = "free_x") +
    scale_fill_brewer(palette = "RdYlGn") +
    # theme(panel.spacing.x = unit(0, "npc")) + # if no spacing preferred between bars
    theme_void()


sub_long2 %>% ggplot(aes(year, Strategy, fill=Count)) + 
    geom_tile(color="white") + scale_x_discrete() 
geom_tile(color="white") + scale_x_discrete() + coord_equal() + theme(axis.ticks.y = element_blank(),panel.background = element_blank(),legend.position = c(0.87, 1.21),legend.background = element_blank(),legend.direction = "horizontal") + scale_fill_gradient(labels = c("Never", "Always"))

```
