---
title: "Transit Cost Project"
author: "Mike Wehinger"
date: "05/01/2021"
output: html_document
---

```{r setup, include=FALSE}


knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, cache = FALSE)

library(tidyverse)
library(leaflet)

transit_cost <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-01-05/transit_cost.csv')

transit_cost <- transit_cost %>% drop_na() %>% filter(country=="AU")


```

# Exploratory Analysis

The data is from a database that spans more than 50 countries and totals more than 11,000 km of urban rail built since the late 1990s.

## Project Timelines

Will focus on Australia and make a nicely designed project time-line chart that will be useful for other purposes.

```{r timeLineChart, echo=FALSE}

library(ggalt)
library(ggforestplot)

#Convert the start and end years into date variable types...

transit_cost$start_year <- as.Date(paste("01/01/",transit_cost$start_year, sep=""), "%d/%m/%Y")

transit_cost$end_year <- as.Date(paste("01/01/",transit_cost$end_year, sep=""), "%d/%m/%Y")

# Chart...

proj <- transit_cost %>% mutate(line = fct_reorder(line, desc(start_year))) %>%
    ggplot(aes(x=start_year, xend=end_year, y=line, group=city)) + 
    geom_dumbbell(size=1.3, color="#007FA9",
                  colour_x = "#007FA9",
                  colour_xend = "#007FA9",
                  size_x = 5,
                  size_xend = 5,
                  dot_guide=FALSE, dot_guide_size=0.25) + 
  labs(x=NULL, y=NULL,
       title="Project Timelines",
       subtitle="Years start to finish",
       caption="Source: https://transitcosts.com/") + 
  theme_classic() + 
  geom_stripes(odd = "#11111111", even = "#00000000") +
  theme(plot.title = element_text(hjust=0.5, face="bold"),
        panel.grid.major.x = element_line(color = "gray", size = .5),
        axis.ticks.x=element_blank(),
        axis.ticks.y=element_blank(),
        axis.line.x = element_blank()
        )

proj

```

## Project Costs

A simple bar chart for cost per KM 

```{r costChart, echo=FALSE}

# A simple bar graph in orange 

transit_cost %>% mutate(line = fct_reorder(line, cost_km_millions)) %>%
  ggplot(aes(x=line)) +
    geom_bar( aes(y=cost_km_millions), stat="identity", size=.1, fill="#e57d09", color="#b85c01", alpha=.99) +
  coord_flip() +
  labs(x=NULL, y= "Costs",
       title="Project Cost Per KM",
       subtitle="AUD $M") + 
  theme_classic() +
  theme(plot.title = element_text(hjust=0.5, face="bold"),
        panel.grid.major.x = element_line(color = "gray", size = .5)
        )

```

## Maps

Put the projects into a leaflet map...

```{r prepData}

#Add Lat & Long for Australia,
transit_cost[transit_cost$line == "Metro City & SW",'lat'] <- -34.18447 
transit_cost[transit_cost$line == "Metro City & SW",'long'] <- 150.78051 

transit_cost[transit_cost$line == "Metro NW",'lat'] <- -33.05082 
transit_cost[transit_cost$line == "Metro NW",'long'] <- 151.16623 

transit_cost[transit_cost$line == "Metro Tunnel",'lat'] <- -37.817187
transit_cost[transit_cost$line == "Metro Tunnel",'long'] <- 144.953335

transit_cost[transit_cost$line == "Forrestfield-Airport Link",'lat'] <- -31.976550
transit_cost[transit_cost$line == "Forrestfield-Airport Link",'long'] <- 115.990174

```


Plot projects on a map

```{r map, echo=FALSE, eval=TRUE}

transit_cost$content <- paste(sep = "<br/>",
  paste("<b>", transit_cost$line, "</b>", sep=""),
  transit_cost$city,
  paste("Length", transit_cost$length)
)


transit_cost %>% 
  filter(country=="AU") %>% 
  leaflet() %>% 
  addProviderTiles("Stamen.TonerLite") %>%
  addPopups(lat = ~lat, lng = ~long, ~content)


map <- transit_cost %>% 
  filter(country=="AU") %>% 
  leaflet() %>% 
  addProviderTiles("Stamen.TonerLite") %>% 
  addMarkers(lat = ~lat, lng = ~long, label = ~line,
             labelOptions = labelOptions(noHide = T,
                                         direction="right",
                                         style = list(
                                           "color" = "midnightblue",
                                           "font-family" = "serif",
                                           "font-style" = "italic",
                                           "box-shadow" = "3px 3px rgba(0,0,0,0.25)",
                                           "font-size" = "12px",
                                           "border-color" = "rgba(0,0,0,0.5)")))

map
```

```{r mapExplore, echo=FALSE, include=FALSE, eval=FALSE}
# Test All maps...

map <- transit_cost %>% leaflet() %>% addMarkers(lng = ~long, lat = ~lat)

map %>% addProviderTiles("Stamen.TonerLite")
map %>% addProviderTiles("Esri.WorldGrayCanvas") 
map %>% addProviderTiles("CartoDB.VoyagerLabelsUnder")
map %>% addProviderTiles("Stamen.Watercolor")
```

## USA Version

Try again, this time for the USA...

```{r USAChart, echo=FALSE}

transit_cost <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-01-05/transit_cost.csv')

transit_cost <- transit_cost %>% drop_na() %>% filter(country=="US") %>% mutate(line = paste(line, "\n", "(", city, ")", sep=""))

#Convert the start and end years into date variable types...

transit_cost$start_year <- as.Date(paste("01/01/",transit_cost$start_year, sep=""), "%d/%m/%Y")

transit_cost$end_year <- as.Date(paste("01/01/",transit_cost$end_year, sep=""), "%d/%m/%Y")

# Chart...

proj <- transit_cost %>% mutate(line = fct_reorder(line, desc(start_year))) %>%
    ggplot(aes(x=start_year, xend=end_year, y=line, group=city)) + 
    geom_dumbbell(size=1.3, color="#007FA9",
                  colour_x = "#007FA9",
                  colour_xend = "#007FA9",
                  size_x = 5,
                  size_xend = 5,
                  dot_guide=FALSE, dot_guide_size=0.25) + 
  labs(x=NULL, y=NULL,
       title="Project Timelines",
       subtitle="Years start to finish",
       caption="Source: https://transitcosts.com/") + 
  theme_classic() + 
  geom_stripes(odd = "#11111111", even = "#00000000") +
  theme(plot.title = element_text(hjust=0.5, face="bold"),
        panel.grid.major.x = element_line(color = "gray", size = .5),
        axis.ticks.x=element_blank(),
        axis.ticks.y=element_blank(),
        axis.line.x = element_blank()
        )

proj

# A simple bar graph in orange 

costs <- transit_cost %>% mutate(line = fct_reorder(line, cost_km_millions)) %>%
  ggplot(aes(x=line)) +
    geom_bar( aes(y=cost_km_millions), stat="identity", size=.1, fill="#e57d09", color="#b85c01", alpha=.99) +
  scale_y_continuous(labels=scales::dollar_format()) + 
  coord_flip() +
  labs(x=NULL, y= "Costs / KM",
       title="Project Cost Per KM",
       subtitle="USD $M") + 
  theme_classic() +
  theme(plot.title = element_text(hjust=0.5, face="bold"),
        panel.grid.major.x = element_line(color = "gray", size = .5)
        )

costs

costs2 <- transit_cost %>% group_by(city) %>% summarize(avgCostPkm = mean(cost_km_millions)) %>% mutate(line = fct_reorder(city, avgCostPkm)) %>%
  ggplot(aes(x=line)) +
    geom_bar( aes(y=avgCostPkm), stat="identity", size=.1, fill="#e57d09", color="#b85c01", alpha=.99) +
  scale_y_continuous(labels=scales::dollar_format()) + 
  coord_flip() +
  labs(x=NULL, y= "Costs / KM",
       title="Average Project Cost Per KM By City",
       subtitle="USD $M") + 
  theme_classic() +
  theme(plot.title = element_text(hjust=0.5, face="bold"),
        panel.grid.major.x = element_line(color = "gray", size = .5)
        )

costs2

# Get Lat & Long for cities...,
transit_cost[transit_cost$city == "Boston",'lat'] <- 42.35866 
transit_cost[transit_cost$city == "Boston",'long'] <- -71.05674 

transit_cost[transit_cost$city == "Los Angeles",'lat'] <- 34.05349 
transit_cost[transit_cost$city == "Los Angeles",'long'] <- -118.24532 

transit_cost[transit_cost$city == "New York",'lat'] <- 40.71455 
transit_cost[transit_cost$city == "New York",'long'] <- -74.00712 

transit_cost[transit_cost$city == "San Francisco",'lat'] <- 37.77986 
transit_cost[transit_cost$city == "San Francisco",'long'] <- -122.42905 

transit_cost[transit_cost$city == "San Jose",'lat'] <- 37.333332 
transit_cost[transit_cost$city == "San Jose",'long'] <- -121.9 

transit_cost[transit_cost$city == "Seattle",'lat'] <- 47.60357 
transit_cost[transit_cost$city == "Seattle",'long'] <- -122.32945 

map <- transit_cost %>%
  leaflet() %>% 
  addProviderTiles("Stamen.TonerLite") %>% 
  addMarkers(lat = ~lat, lng = ~long, label = ~line,
             labelOptions = labelOptions(noHide = T,
                                         direction="right",
                                         style = list(
                                           "color" = "midnightblue",
                                           "font-family" = "serif",
                                           "font-style" = "italic",
                                           "box-shadow" = "3px 3px rgba(0,0,0,0.25)",
                                           "font-size" = "12px",
                                           "border-color" = "rgba(0,0,0,0.5)")))

map

```