---
title: "NCAA Women's Basketball Tournament Analysis"
author: "Mike Wehinger"
date: "12 October 2020"
always_allow_html: true
output:
  md_document
---

```{r setup, include=FALSE, echo=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotly)
### Load & clean data (the first time) and save as a csv for future loads. 
  # raw_df <- read_csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/ncaa-womens-basketball-tournament/ncaa-womens-basketball-tournament-history.csv")

  # clean_tourn <- raw_df %>% 
  #   janitor::clean_names() %>% 
  #   mutate(across(c(seed, conf_w:conf_percent, full_percent), parse_number))

  # clean_tourn %>% 
  #   write_csv("tourny.csv")

# Get Data
tourny <- read_csv("tourny.csv")

#Convert Finishing round indicator to a number 
tourny$finRound <- case_when(
  tourny$tourney_finish == "Champ" ~ 7,
  tourny$tourney_finish == "N2nd" ~ 6,
  tourny$tourney_finish == "NSF" ~ 5,
  tourny$tourney_finish == "RF" ~ 4,
  tourny$tourney_finish == "RSF" ~ 3,
  tourny$tourney_finish == "2nd" ~ 2,
  tourny$tourney_finish == "1st" ~ 1,)

# Remove rows where there is no finih round data
tourny <- tourny %>% drop_na(finRound) %>% drop_na(seed)

# Create a reference table for counts of teams that finish in each round. This will be used for calculating probabilities...
RoundCounts <- tourny %>% group_by(finRound) %>% summarize(RoundFreq=n())

#Calculate probabilities of each seed making it to their respective final rounds
scatterData <- tourny %>% group_by(seed, finRound) %>% summarize(freq = n()) %>% arrange(finRound) %>% left_join(RoundCounts) %>% mutate(prob = (round((freq / RoundFreq)*100,1)))

#Format the probability for the chart 
scatterData[scatterData$prob >=1,]$prob <- round(scatterData[scatterData$prob >=1,]$prob,0)

# Highlight teams that have overcome low odds aka cinderellas
scatterData$cinderella <- case_when(scatterData$prob < 4 & scatterData$finRound > 2 ~ "Cinderella", TRUE ~ "")

#highlight certain teams
a <- list(
    x = tourny[tourny$seed == 11 & tourny$finRound == 4,]$finRound,
    y = tourny[tourny$seed == 11 & tourny$finRound == 4,]$seed,
    text = paste(tourny[tourny$seed == 11 & tourny$finRound == 4,]$school,", ",tourny[tourny$seed == 11 & tourny$finRound == 4,]$year,sep="" ),
    xref = "x",
    yref = "y",
    showarrow = TRUE,
    arrowhead = 4,
    ax = 20,
    ay = -40
)

b <- list(
    x = tourny[tourny$seed == 9 & tourny$finRound == 5,]$finRound,
    y = tourny[tourny$seed == 9 & tourny$finRound == 5,]$seed,
    text = paste(tourny[tourny$seed == 9 & tourny$finRound == 5,]$school,", ",tourny[tourny$seed == 9 & tourny$finRound == 5,]$year,sep="" ),
    xref = "x",
    yref = "y",
    showarrow = TRUE,
    arrowhead = 4,
    ax = 20,
    ay = -40
)

c <- list(
    x = tourny[tourny$seed == 5 & tourny$finRound == 6,]$finRound,
    y = tourny[tourny$seed == 5 & tourny$finRound == 6,]$seed,
    text = paste(tourny[tourny$seed == 5 & tourny$finRound == 6,]$school,", ",tourny[tourny$seed == 5 & tourny$finRound == 6,]$year,sep="" ),
    xref = "x",
    yref = "y",
    showarrow = TRUE,
    arrowhead = 4,
    ax = 20,
    ay = -40
)

d <- list(
    x = scatterData$finRound,
    y = scatterData$seed,
    text = scatterData$prob,
    xref = "x",
    yref = "y",
    textfont = list(color = '#000000', size = 3),
    showarrow = F,
    arrowhead = 4,
    ax = 0,
    ay = 0
)

fig <- scatterData %>% plot_ly(x = ~finRound, y = ~seed,
                        colors = c("blue", "coral2"), 
                        color = ~cinderella,
                        marker = list(symbol = "square-open", size=18, line = list(color = 'rgba(10, 10, 10, .5)',width = 1)), 
                        text=paste("Seed: ", scatterData$seed, "\n", "Round: ", scatterData$finRound, "\n", "prob: ", scatterData$prob,"%", sep=""),hoverinfo = 'text',showlegend = FALSE)

fig <- fig %>% layout(annotations = a) %>% layout(annotations = b)%>% layout(annotations = c)%>% layout(annotations = d)%>% layout(title = 'NCAAW Tournament Probabilities',xaxis = list(title = paste("Final Round", "\n","( 1=1st Round, 7 = Championship)", sep="")),yaxis = list(title = "Seed"))

```

## Seed probabilities

The chart below shows the probabilities of a team finishing in each round based on their seed. For example,

1. There is a 76% chance that a number one seeded team will when the championship (round 7)

2. There is a .1% chance that a number one seeded team will exit the tournament in round one. 

The pink highlighted points show the Cinderella teams. These are high seeded teams that make it deep into the tournament. These scenarios have a less than 4% chance of occurring.   


```{r SeedChart, echo=FALSE, warning=FALSE,message = FALSE}
fig


```

