---
title: "Plastic"
author: "Mike Wehinger"
date: "27/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(kableExtra)
library(scales)

#plastics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-01-26/plastics.csv')

plastics <- readr::read_csv('plastics.csv')

# Resources
# https://ggraph.data-imaginist.com/reference/geom_node_voronoi.html
# https://ggraph.data-imaginist.com/articles/Edges.html
# https://www.data-to-viz.com/graph/dendrogram.html
# http://exts.ggplot2.tidyverse.org/ggraph.html
# https://www.data-imaginist.com/2017/ggraph-introduction-edges/
# https://www.r-graph-gallery.com/334-basic-dendrogram-with-ggraph.html


```

## Explore Data

Will look at change in counts or proportion of plastic types over time. 


#### Country



Look at data collection by country and remove inconsistencies and counts < 10

```{r country}

# There is a lot of variation in data collection from year to year... 

country <- plastics %>% 
        group_by(country, year) %>% 
        summarize(count = n()) %>% 
        filter(count>10) %>% 
        pivot_wider(names_from = year, values_from = count) %>% 
        drop_na() %>% 
        mutate(pChange = (sum(`2020`) - sum(`2019`)) / sum(`2019`)) %>%
        arrange(desc(pChange))

country

country %>% ggplot(aes(pChange)) + geom_histogram() + scale_x_continuous(labels = scales::percent)

```

#### Brand

Look at data by brand 

```{r brand}

# There are a lot of brands w/ small counts. There may be a need to clean brand names...

brand <- plastics %>% group_by(parent_company, year) %>% summarize(count = n()) %>% pivot_wider(names_from = year, values_from = count) %>% drop_na() %>% mutate(pChange = (sum(`2020`) - sum(`2019`)) / sum(`2019`)) %>% arrange(parent_company)

brand %>% filter(`2020` > 5)

brand %>% ggplot(aes(pChange)) + geom_histogram() + scale_x_continuous(labels = scales::percent)

brand %>% arrange(desc(`2020`)) %>% filter(`2020` > 10) %>% pivot_longer(c(`2019`,`2020`), names_to = "year", values_to = "count") %>% ggplot(aes(x=parent_company, y=count, fill=year)) + geom_col(position = position_dodge()) + coord_flip()

```

#### Plastic Type

Look at plastic types, quwestions, 

relationship between country and brand
proportion changes between years
proportion changes between years and country
```{r plastic}

pTypes <- plastics %>% drop_na %>% group_by(year) %>% summarize(hdpe = (sum(hdpe) / sum(grand_total)), ldpe = (sum(ldpe) / sum(grand_total)), o = (sum(o) / sum(grand_total)), pet = (sum(pet) / sum(grand_total)), pp = (sum(pp) / sum(grand_total)), ps = (sum(ps) / sum(grand_total)), pvc = (sum(pvc) / sum(grand_total)))

pTypes %>% pivot_longer(hdpe:pvc) %>% ggplot(aes(x=name, y=value, fill=as.factor(year))) + geom_col(position = position_dodge()) + coord_flip()

pTypesCountry <- plastics %>% drop_na %>% group_by(country, year) %>% summarize(hdpe = (sum(hdpe) / sum(grand_total)), ldpe = (sum(ldpe) / sum(grand_total)), o = (sum(o) / sum(grand_total)), pet = (sum(pet) / sum(grand_total)), pp = (sum(pp) / sum(grand_total)), ps = (sum(ps) / sum(grand_total)), pvc = (sum(pvc) / sum(grand_total)))

# Look at proportions of plastic types by country and year...


```

#### Events & Volunteers


```{r events}

events <- plastics %>% group_by(year, country) %>% summarize(itemsFound = sum(grand_total), volunteers = mean(volunteers), events = mean(num_events)) %>% mutate(itemsPvolunteer = itemsFound/volunteers, volunteersPevent = volunteers/events)


```

```{r test}
library(ggraph)
library(igraph)

data=data.frame(
  level1="CEO",
  level2=c( rep("boss1",4), rep("boss2",4)),
  level3=paste0("mister_", letters[1:8])
)
 
# transform it to a edge list!
edges_level1_2 = data %>% select(level1, level2) %>% unique %>% rename(from=level1, to=level2)
edges_level2_3 = data %>% select(level2, level3) %>% unique %>% rename(from=level2, to=level3)
edge_list=rbind(edges_level1_2, edges_level2_3)
 
# Now we can plot that
mygraph <- graph_from_data_frame( edge_list )
ggraph(mygraph, layout = 'dendrogram', circular = FALSE) + 
  geom_edge_diagonal() +
  geom_node_point(color="#69b3a2", size=3) +
  geom_node_text(
    aes(  label=c("CEO", "Manager", "Manager", LETTERS[8:1]) ), 
    hjust=c(1,0.5, 0.5, rep(0,8)), 
    nudge_y = c(-.02, 0, 0, rep(.02,8)),
    nudge_x = c(0, .3, .3, rep(0,8))
  ) +
  theme_void() +
  coord_flip() +
  scale_y_reverse() 

```


```{r test2}
library(ggraph)
library(igraph)

data=data.frame(
  level1="US",
  level2=c( rep("hdpe",5), rep("ldpe",5),rep("o",5), rep("pet",5),rep("pp",5), rep("ps",5), rep("pvc",5)),
  level3=paste0("mister_", 1:35)
)
 
# transform it to a edge list!
edges_level1_2 = data %>% select(level1, level2) %>% unique %>% rename(from=level1, to=level2)
edges_level2_3 = data %>% select(level2, level3) %>% unique %>% rename(from=level2, to=level3)
edge_list=rbind(edges_level1_2, edges_level2_3)
 
# Now we can plot that
mygraph <- graph_from_data_frame( edge_list )
ggraph(mygraph, layout = 'dendrogram', circular = FALSE) + 
  geom_edge_diagonal() +
  geom_node_point(color="#69b3a2", size=3) +
  geom_node_text(
    aes(  
            label=c("USA", "hdpe", "ldpe", "o", "pet", "pp", "ps", "pvc", paste("co",1:35, sep="_"))
            ), 
    hjust=c(1,rep(.5,7), rep(0,35)), 
    nudge_y = c(-.02,rep(0,7), rep(.02,35)),
    nudge_x = c(0,rep(1,7), rep(0,35))
  ) +
  theme_void() +
  coord_flip() +
  scale_y_reverse() 

#possible unicode symbols...
# \u2297 ⊗
# \u2022 •
# \u220E ∎
# \u2587 ▇
# \u25C9 ◉
# \u25CF ●
# Good resource: https://www.htmlsymbols.xyz/unicode?page=35


```
